\section{Foundational Features for Integrating Data and Computation}

\subsection{Unknown Values}

We introduce a new feature for unknown values $?_A$ of type $A$.
These represent meaningful but unknown values.

This is similar to exceptions, undefined values, or option types except that there is no implication that the original expression is invalid.
Instead, the original expression is well-typed and well-defined, but its concrete value is not known.
(This is similar to Scala's $???$.)

$?_A$ can only occur in definitions.
Definition expansion fails if simplification would eventually encounters a term of the form $?_A$.

Challenges:
\begin{itemize}
 \item $?_A$ does not respect congruence.
 \item Where exactly can $?_A$ occur?
 \item How can we predict whether definition expansion should be allowed? For example, might contain $?_A$ but never fail for the argument at hand.
\end{itemize}

\subsection{Persistent Memoization of Functions}

We allow any function $f:A \to B$ to be coupled with a table of values, i.e., pairs $(a,b)\in A\times B$ such that $f(a)=b$.

These can be stored in a database.

Virtual theories and persistent memoization are the two key concepts to combine mathematical computation and data.
They correspond to the efficient storage of sets and maps.

Challenges:
\begin{itemize}
 \item What is the content representation of a tabled function?
 \item How are functions associated with databases?
 \item How is lookup done if the argument is not normalized?
 \item If we think of $\to$ as an abstract class (like in Scala), we can create memoized functions as special cases.
 Then memoization could be a more general concept that allows storing arbitrary lists of objects.
\end{itemize}

\subsection{Exceptional Values}

This is similar to memoization of functions except that the given values supersede the result of the function.

This can be represented as a record $(exc: A \to bool, f: A \to B, V: Set[A*B])$.
Its meaning is the function which maps $x:A$ to the lookup in $V$ if $exc(x)$ and to $f(x)$ otherwise.
The set $V$ is then memoized.